---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linking dimensions of data on global marine animal diversity

This is the code for reproducing the figures and analysis reported in Webb & Vanhoorne _Linking dimensions of data on global marine animal diversity_, Phil Trans R Soc B, doi: 10.1098/rstb.2019.0445

First, load required packages (see bottom of this script for installed versions used):
```{r, message = FALSE}
library(tidyverse) # For reading, manipulating and plotting data
library(pscl) # For fitting hurdle models
library(broom) # For tidying model outputs
library(patchwork) # For multi-panel figure layouts
library(ggmosaic) # For mosaic plots
library(ggbeeswarm) # For beeswarm-style plots
```

The dataset used in this analysis is available under a Creative Commons Attribution 4.0 International License in the Marine Data Archive (MDA). The preferred route to access the data is via IMIS (Integrated Marine Inforamtion System), which provides full meta-data. The IMIS record for our dataset is here: https://doi.org/10.14284/417 The code below uses the direct link to MDA to download and unzip the two component csv files into the data folder of this project directory: the full dataset and a second csv which provides a full description of each variable.

```{r, message = FALSE, warning = FALSE}
# download zipped data from MDA
curl::curl_download(url = "https://mda.vliz.be/download.php?file=VLIZ_00000106_5f1ff6c0cf0a1110164358",
                    destfile = "data/worms_animals_links.zip")
# unzip
unzip("data/worms_animals_links.zip", exdir = "data")
```

The main dataset can now be read in:
```{r, message = FALSE, warning = FALSE}
worms_valid <- readr::read_csv("data/worms_animals_links.csv")
```

This contains `r nrow(worms_valid)` observations of `r ncol(worms_valid)` variables. Each observation is a valid marine animal species. Variables are described fully in the meta data here:
```{r, message = FALSE, warning = FALSE}
worms_valid_meta <- readr::read_csv("data/meta.csv")
```
For each variable, this gives its class (`class`), the number of non-NA observations (`n_obs`), the number of missing (NA) observations (`n_miss`), the percent missing observations (`pct_miss`), a description of the variables (`Description`), and relevant notes (`Notes`). For info: the missing value scores were generated using `naniar::miss_var_summary`.

## Visualising data availability across phyla and functional groups
Next, we create a phylum-level summary dataset, which summarises the number of species per phylum, together with the number of species with records in OBIS, GenBank, neither, or both. It also gives phylum-level median number of non-zero OBIS and GenBank records:
```{r, message = FALSE}
phylum_obis_gen_summary <- worms_valid %>% group_by(phylum) %>% 
  summarise(
    n_sp = n(),
    n_obis = sum(!is.na(OBIS) & is.na(GenBank_nuc)),
    n_obis_gen = sum(!is.na(OBIS) & !is.na(GenBank_nuc)),
    n_gen = sum(is.na(OBIS) & !is.na(GenBank_nuc)),
    n_neither = sum(is.na(OBIS) & is.na(GenBank_nuc)),
    med_obis = median(OBIS, na.rm = TRUE),
    med_gen = median(GenBank_nuc, na.rm = TRUE)
  ) %>% 
  rownames_to_column() %>% 
  rename(phylum_id = rowname) %>% 
  mutate(phylum_id = as.numeric(phylum_id))

```
We then create a long version of this summary data frame ready for plotting, adding a scaling factor so that the number of species per phylum is rescaled onto 0.25-1, for nice plotting of bar widths:
```{r}
phylum_obis_gen_long <- phylum_obis_gen_summary %>%
  pivot_longer(
    cols = n_obis:n_neither,
    names_to = "data_source",
    values_to = "n_species"
  ) %>% 
  mutate(data_source = ordered(data_source, c("n_neither", "n_gen", "n_obis_gen", "n_obis")),
         p_species = n_species / n_sp,
         scaled_sp = ((n_sp / 76448) + 0.25))

```
Create a colour palette for plotting this chart of proportion of species per phylum with data in each database:
```{r}
data_source_pal <- tibble(data_source = c("n_neither", "n_gen", "n_obis_gen", "n_obis"),
                          ds_colours = c("#FCFAF1", "#DB565D", "#633231", "#FACCAD"),
                          ds_labels = c("no data", "GenBank only", "OBIS and GenBank", "OBIS only")
                          )
```
And now create the plot (Figure 1A):
```{r}
p_species_data_source <- ggplot(phylum_obis_gen_long,
                                aes(x = phylum, y = p_species,
                                    fill = data_source, width = scaled_sp)) +
  geom_bar(stat = "identity", position = "stack", alpha = 0.8, colour = "black", size = 0.2) +
  scale_fill_manual(name = "P(species with data)",
                    values = pull(data_source_pal, ds_colours),
                    labels = pull(data_source_pal, ds_labels)) +
  labs(x = "", y = "", tag = "A") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
  scale_y_continuous(breaks=c(0, 0.5, 1)) +
  geom_hline(yintercept = c(0, 0.5, 1), size = 0.25)

```
To create Figure 1 B and C, we use this palette for functional groups:
```{r}
functional_group_palette <- tibble(
  functional_group = c("benthos", "zooplankton", "nekton",
                       "fish", "mammals", "birds", "reptiles",
                       "other/unknown"),
  fg_pal = c("#077893", "#FF9465", "#00AE91",
             "#45BEC6", "#9F4440", "#FFD37B", "#EE9BE4",
             "#E3DECA")
)
```

Then generate the figures, plotting the number of OBIS records and then GenBank nucleotides for each species (excluding 0 records) in each phylum:
```{r, warning = FALSE}
obis_by_phylum <- ggplot(worms_valid, aes(x = phylum, y = log10(OBIS))) +
  geom_jitter(aes(colour = functional_group), size = 1/3, alpha = 3/5, na.rm = TRUE) +
  scale_colour_manual(name = "functional group",
                      values = pull(functional_group_palette, fg_pal),
                      labels = pull(functional_group_palette, functional_group)) +
  labs(x = "", y = expression(log[10]*"(OBIS records)"), tag = "B") +
  theme_minimal() +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), legend.position = "none") +
  geom_boxplot(na.rm = TRUE, fill = NA, size = 0.25, outlier.shape = NA,
               fatten = 0, varwidth = TRUE) +
  geom_point(data = phylum_obis_gen_summary, aes(x = phylum, y = log10(med_obis)),
             shape = 4, size = 1.5, stroke = 1.1)

genbank_by_phylum <- ggplot(worms_valid, aes(x = phylum, y = log10(GenBank_nuc))) +
    geom_jitter(aes(colour = functional_group), size = 1/3, alpha = 3/5, na.rm = TRUE) +
    scale_colour_manual(name = "functional group",
                        values = pull(functional_group_palette, fg_pal),
                        labels = pull(functional_group_palette, functional_group)) +
    labs(x = "", y = expression(log[10]*"(GenBank nucleotides)"), tag = "C") +
    theme_minimal() +
    theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size = 5))) +
    geom_boxplot(na.rm = TRUE, fill = NA, size = 0.25, outlier.shape = NA,
                 fatten = 0, varwidth = TRUE) +
    geom_point(data = phylum_obis_gen_summary, aes(x = phylum, y = log10(med_gen)),
               shape = 4, size = 1.5, stroke = 1.1)
```
We use the `patchwork` package to assemble these three panels:
```{r, warning = FALSE}
(linking_fig1 <- genbank_by_phylum / obis_by_phylum / p_species_data_source +
  plot_layout(heights = c(2, 2, 1))
)
```
  
This can then be saved in your required format:
```{r, warning = FALSE}
ggsave("figures/webb_philtrans_fig1.pdf", linking_fig1, width = 10, height = 8, units = "in")  
```

## Data availablity across fish habitat groups
For fish we also explored data availability across habitat groups derived from FishBase, using the `rfishbase` package to add habitat affinities to our dataset (stored in `functional_group2`). Filter the dataset to fish only, and order the categories:
```{r}
fish <- worms_valid %>% filter(functional_group == "fish") %>%
  mutate(functional_group2 = ifelse(
    functional_group2 == "fish_unspecified_habitat", "habitat unspecified", functional_group2)) %>% 
  mutate(functional_group2 = ordered(functional_group2,
                                            c("bathydemersal", "bathypelagic",
                                            "benthopelagic", "demersal",
                                            "pelagic-oceanic", "pelagic-neritic",
                                            "reef-associated", "habitat unspecified")))
```
Then create a summary of data availablity across these groups:
```{r, message = FALSE}
fish_obis_gen_summary <- fish %>% group_by(functional_group2) %>% 
  summarise(
    n_sp = n(),
    n_obis = sum(!is.na(OBIS) & is.na(GenBank_nuc)),
    n_obis_gen = sum(!is.na(OBIS) & !is.na(GenBank_nuc)),
    n_gen = sum(is.na(OBIS) & !is.na(GenBank_nuc)),
    n_neither = sum(is.na(OBIS) & is.na(GenBank_nuc)),
    med_obis = median(OBIS, na.rm = TRUE),
    med_gen = median(GenBank_nuc, na.rm = TRUE)
  ) %>% 
  rownames_to_column() %>% 
  rename(habitat_id = rowname)
```
Make this long for plotting:
```{r}
fish_obis_gen_long <- fish_obis_gen_summary %>%
  pivot_longer(
    cols = n_obis:n_neither,
    names_to = "data_source",
    values_to = "n_species"
  ) %>% 
  mutate(data_source = ordered(data_source, c("n_neither", "n_gen", "n_obis_gen", "n_obis")),
         p_species = n_species / n_sp,
         scaled_sp = (n_sp / (6248/0.75)) + 0.25)
```
Create the first panel of the figure - proportion of species in each habiat class with data in each database:
```{r}
p_species_data_source_fish <- ggplot(fish_obis_gen_long,
                                      aes(x = functional_group2, y = p_species,
                                          fill = data_source, width = scaled_sp)) +
    geom_bar(stat = "identity", position = "stack", alpha = 0.8, colour = "black", size = 0.2) +
    scale_fill_manual(name = "P(species with data)",
                      values = pull(data_source_pal, ds_colours),
                      labels = pull(data_source_pal, ds_labels)) +
    labs(x = "", y = "", tag = "A") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
    scale_y_continuous(breaks=c(0, 0.5, 1)) +
    geom_hline(yintercept = c(0, 0.5, 1), size = 0.25)
```
For the remaining panels, use this palette for fish habitats:
```{r}
fish_habitat_palette <- tibble(
  habitat = c("bathydemersal", "bathypelagic",
                        "benthopelagic", "demersal",
                        "pelagic-oceanic", "pelagic-neritic",
                        "reef-associated", "habitat unspecified"),
  fg_pal = c("#003E51", "#2B4D50", "#3C3768", "#486D87",
             "#007DBA", "#2E94AC", "#9DB6B5", "#C5B9AC")
)
```
Plot number of OBIS records per species by habitat class:
```{r, warning = FALSE}
obis_by_fish_fg <- 
  ggplot(fish, aes(x = functional_group2, y = log10(OBIS))) +
  geom_quasirandom(aes(colour = functional_group2), size = 1/3, alpha = 3/5, na.rm = TRUE) +
  scale_colour_manual(name = "habitat",
                      values = pull(fish_habitat_palette, fg_pal),
                      labels = pull(fish_habitat_palette, habitat)) +
  labs(x = "", y = expression(log[10]*"(OBIS records)"), tag = "B") +
  theme_minimal() +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), legend.position = "none") +
  geom_boxplot(na.rm = TRUE, fill = NA, size = 0.25, outlier.shape = NA,
               fatten = 0, varwidth = TRUE) +
  geom_point(data = fish_obis_gen_summary, aes(x = functional_group2, y = log10(med_obis)),
             shape = 4, size = 1.5, stroke = 1.1)

```
And the same for GenBank nucleotides:
```{r, warning = FALSE}
genbank_by_fish_fg <- 
  ggplot(fish, aes(x = functional_group2, y = log10(GenBank_nuc))) +
  geom_quasirandom(aes(colour = functional_group2), size = 1/3, alpha = 3/5, na.rm = TRUE) +
  scale_colour_manual(name = "habitat",
                      values = pull(fish_habitat_palette, fg_pal),
                      labels = pull(fish_habitat_palette, habitat)) +
  labs(x = "", y = expression(log[10]*"(Genbank nucleotides)"), tag = "C") +
  theme_minimal() +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
  guides(colour = guide_legend(override.aes = list(size = 5))) +
  geom_boxplot(na.rm = TRUE, fill = NA, size = 0.25, outlier.shape = NA,
               fatten = 0, varwidth = TRUE) +
  geom_point(data = fish_obis_gen_summary, aes(x = functional_group2, y = log10(med_gen)),
             shape = 4, size = 1.5, stroke = 1.1)

```
Assemble Figure S1:
```{r, warning = FALSE}
(linking_fig1_fish_habs <- genbank_by_fish_fg / obis_by_fish_fg / p_species_data_source_fish +
    plot_layout(heights = c(2, 2, 1))
)
```
  
And save to file:
```{r, warning = FALSE}
ggsave("figures/webb_philtrans_figS1.pdf", linking_fig1_fish_habs,
       width = 10, height = 8, units = "in")
```

## Modelling data availability across functional groups
Because of the large number of zeros in our data, we use a hurdle modelling approach to model data availability across functional groups. First, we create a couple of convenience variables to aid modelling:

```{r}
worms_valid <- worms_valid %>% 
  mutate(n_obis = ifelse(is.na(OBIS), 0, OBIS),
         n_gb = ifelse(is.na(GenBank_nuc), 0, GenBank_nuc),
         fg = factor(functional_group))
```
We then run a single hurdle model, modelling zeros as binomial and non-zero counts as negative binomial, using `pscl::hurdle`:

```{r}
fm_obis_hurdle <- hurdle(n_obis ~ fg - 1, data = worms_valid,
                         dist = "negbin", zero.dist = "binomial")
```
To derive estimated non-zero counts for each functional group, we first get P(non-zero count) for each functional group:
```{r}
zero_coefs <- coef(fm_obis_hurdle)[which(str_detect(names(coef(fm_obis_hurdle)), "zero"))]
p_nonzero <- exp(zero_coefs) / (exp(zero_coefs) + 1)
```
Mean count (including zeros) is just the response prediction:
```{r}
mean_count <- predict(fm_obis_hurdle, type = "response",
                      newdata = data.frame(fg = levels(worms_valid$fg)))
```
So predicted counts for non-zero species are:
```{r}
count_given_present <- mean_count / p_nonzero
```
Create a simple summary table for functional groups:
```{r, message = FALSE}
fg_summary <- worms_valid %>% 
  group_by(functional_group) %>% 
  summarise(n_sp = n(), n_obis = sum(in_obis), n_genb = sum(in_genb)) %>% 
  mutate(functional_group = ordered(functional_group,
                                    c("benthos", "zooplankton", "nekton", "fish",
                                      "mammals", "birds", "reptiles", "other/unknown")))
```
For convenience we run the zero model as a separate binomial GLM to facilitate calculation of CIs:
```{r}
fm_in_obis <- glm(in_obis ~ functional_group - 1, family = binomial, data = worms_valid)
```
Now create a tidy version of the coefficients:
```{r}
fm_in_obis_tidy <- tidy(fm_in_obis, conf.int = TRUE, exponentiate = TRUE) %>% 
  bind_cols(fg_summary)
```
Get empirical means and bootstrap CIs for the non-zero counts. This function will do this based on a sample of 90 (given 96 species in the smallest groups):
```{r}
get_fg_means <- function(dat, i = 90){
  dat %>% 
    group_by(functional_group) %>% 
    sample_n(i, replace = T) %>% 
    summarise(mean_obis = mean(OBIS, na.rm = TRUE), .groups = "keep")
}
```
Run this 10,000 times:
```{r}
bootstrap_fg_means <- purrr::map_dfr(1:10000, function(x) get_fg_means(worms_valid))
```
Get number of species in OBIS for each functional group:
```{r}
n_in_obis <- worms_valid %>% filter(!is.na(OBIS)) %>% count(functional_group)
```
Get dataframe of mean from data and from model:
```{r, message = FALSE}
mean_counts <- worms_valid %>%
  filter(in_obis == 1) %>%
  group_by(functional_group) %>%
  summarise(mean_from_data = mean(OBIS)) %>% 
  mutate(est_from_model = count_given_present)
```
Create a summary data frame:
```{r, message = FALSE}
bootstrap_obis_summary <- bootstrap_fg_means %>%
  filter(!is.na(mean_obis)) %>% 
  summarise(obis_non_zero = mean(mean_obis),
            low_ci = quantile(mean_obis, 0.025),
            high_ci = quantile(mean_obis, 0.975)) %>% 
  left_join(mean_counts, by = "functional_group") %>% 
  left_join(n_in_obis, by = "functional_group") %>% 
  rename(n_obis_sp = n) %>% 
  mutate(functional_group = ordered(functional_group,
                                    c("benthos", "zooplankton", "nekton", "fish",
                                      "mammals", "birds", "reptiles", "other/unknown")))
```
Now create the figure. First, zero coefficients:
```{r, warning = FALSE}
in_obis_binomial <- ggplot(fm_in_obis_tidy, aes(x = functional_group, y = estimate,
                                                colour = functional_group)) +
    geom_point(aes(size = log10(n_sp))) +
    geom_segment(aes(x = functional_group, xend = functional_group,
                     y = conf.low, yend = conf.high)) +
    scale_colour_manual(values = pull(functional_group_palette, fg_pal)) +
    labs(x = "", y = "Binomial coefficient (OBIS)", tag = "A") +
    theme_minimal() +
    theme(legend.position = "none") +
    coord_flip()
```
Now for the zero-truncated means:
```{r, warning = FALSE}
obis_n_empirical <- ggplot(bootstrap_obis_summary,
                            aes(x = functional_group, y = log10(mean_from_data),
                                colour = functional_group)) +
    geom_point(aes(size = log10(n_obis_sp))) +
    geom_segment(aes(x = functional_group, xend = functional_group,
                     y = log10(low_ci), yend = log10(high_ci))) +
    geom_point(aes(y = log10(est_from_model)), shape = 4, colour = "black") +
    scale_colour_manual(values = pull(functional_group_palette, fg_pal)) +
    labs(x = "", y = expression(log[10]*"(zero-truncated mean OBIS records)"), tag = "B") +
    theme_minimal() +
    theme(legend.position = "none") +
    coord_flip()
```
Now run the same process for GenBank:
```{r, message = FALSE}
fm_gb_hurdle <- hurdle(n_gb ~ fg - 1, data = worms_valid,
                         dist = "negbin", zero.dist = "binomial")
# work out estimated non-zero counts, first get P(non-zero count) for each functional group
zero_coefs_gb <- coef(fm_gb_hurdle)[which(str_detect(names(coef(fm_gb_hurdle)), "zero"))]
p_nonzero_gb <- exp(zero_coefs_gb) / (exp(zero_coefs_gb) + 1)
# mean count is just the response predictiosn
mean_count_gb <- predict(fm_gb_hurdle, type = "response", newdata = data.frame(fg = levels(worms_valid$fg)))
# so predicted counts for non-zero species are:
count_given_present_gb <- mean_count_gb / p_nonzero_gb

# run the zero model as a separate glm to get CIs:
fm_in_gb <- glm(in_genb ~ functional_group - 1, family = binomial, data = worms_valid)
# now create a tidy version of the coefficients:
fm_in_gb_tidy <- tidy(fm_in_gb, conf.int = TRUE, exponentiate = TRUE) %>% 
  bind_cols(fg_summary)

# get empirical means and bootstrap CIs for the non-zero counts
get_fg_means <- function(dat, i = 75){
  dat %>% 
    group_by(functional_group) %>% 
    sample_n(i, replace = T) %>% 
    summarise(mean_gb = mean(GenBank_nuc, na.rm = TRUE), .groups = "keep")
}

# run this 10,000 times
bootstrap_fg_means_gb <- purrr::map_dfr(1:10000, function(x) get_fg_means(worms_valid))

# get number of species in GenBank for each functional group:
n_in_gb <- worms_valid %>% filter(in_genb == 1) %>% count(functional_group)

# get dataframe of mean from data and from model
mean_counts_gb <- worms_valid %>%
  filter(in_genb == 1) %>%
  group_by(functional_group) %>%
  summarise(mean_from_data = mean(GenBank_nuc)) %>% 
  mutate(est_from_model = count_given_present_gb)

# create a summary data frame
bootstrap_gb_summary <- bootstrap_fg_means_gb %>%
  filter(!is.na(mean_gb)) %>% 
  summarise(gb_non_zero = mean(mean_gb),
            low_ci = quantile(mean_gb, 0.025),
            high_ci = quantile(mean_gb, 0.975)) %>% 
  left_join(mean_counts_gb, by = "functional_group") %>% 
  left_join(n_in_gb, by = "functional_group") %>% 
  rename(n_gb_sp = n) %>% 
  mutate(functional_group = ordered(functional_group,
                                    c("benthos", "zooplankton", "nekton", "fish",
                                      "mammals", "birds", "reptiles", "other/unknown")))
```
Now create the figure. First for zero coefficients:
```{r, warning = FALSE}
in_gb_binomial <- ggplot(fm_in_gb_tidy, aes(x = functional_group, y = estimate,
                                            colour = functional_group)) +
    geom_point(aes(size = log10(n_sp))) +
    geom_segment(aes(x = functional_group, xend = functional_group,
                     y = conf.low, yend = conf.high)) +
    scale_colour_manual(values = pull(functional_group_palette, fg_pal)) +
    labs(x = "", y = "Binomial coefficient (GenBank)", tag = "C") +
    theme_minimal() +
    theme(legend.position = "none") +
    coord_flip()
```
Then for zero-truncated means:
```{r, warning = FALSE}
gb_n_empirical <- ggplot(bootstrap_gb_summary,
                            aes(x = functional_group, y = log10(mean_from_data),
                                colour = functional_group)) +
    geom_point(aes(size = log10(n_gb_sp))) +
    geom_segment(aes(x = functional_group, xend = functional_group,
                     y = log10(low_ci), yend = log10(high_ci))) +
    geom_point(aes(y = log10(est_from_model)), shape = 4, colour = "black") +
    scale_colour_manual(values = pull(functional_group_palette, fg_pal)) +
    labs(x = "", y = expression(log[10]*"(zero-truncated mean GenBank nucleotides)"), tag = "D") +
    theme_minimal() +
    theme(legend.position = "none") +
    coord_flip()
```
Use `patchwork` to assemble Figure 2:
```{r}
(coef_figs <- (in_obis_binomial | obis_n_empirical) / (in_gb_binomial | gb_n_empirical))
```
  
And save to file:
```{r}
ggsave("figures/webb_philtrans_fig2.pdf", coef_figs, width = 10, height = 8, units = "in")
```

## Relationship between OBIS and GenBank data availability
We use mosaic plots to visualise the contingency between OBIS records and GenBank nucleotides. First, we create log-10 categories for both OBIS and GenBank:
```{r}
worms_valid <- worms_valid %>% mutate(
  obis_cat = as.numeric(cut(log10(OBIS),
                              breaks = c(-1:5, 7),
                              labels = 10^(0:6), include.lowest = TRUE, right = TRUE)),
  gen_cat = as.numeric(cut(log10(GenBank_nuc),
                            breaks = c(-1:5, 7),
                            labels = 10^(0:6), include.lowest = TRUE, right = TRUE))
  ) %>% 
  mutate(
    obis_cat = case_when(
      is.na(obis_cat) ~ 0,
      TRUE ~ 10^(obis_cat - 1)),
    gen_cat = case_when(
      is.na(gen_cat) ~ 0,
      TRUE ~ 10^(gen_cat - 1))
    ) %>% 
  mutate(obis_cat = as.factor(obis_cat),
         gen_cat = as.factor(gen_cat))
```
Create a convenient vector of scale labels:
```{r}
p10_scale <- c(0, 1, 10, expression(10^2), expression(10^3),
               expression(10^4), expression(10^5), expression(10^7)
)
```
Mosaic plot of all species:
```{r}
(obis_gen_all <- ggplot(data = worms_valid) +
  geom_mosaic(aes(x = product(obis_cat), fill = gen_cat), offset = 0.001) +
  scale_fill_viridis_d(name = "GenBank\nnucleotides",
                       labels = p10_scale) +
  scale_x_productlist(labels = c(p10_scale[1:4], "", expression(10^3*"-"*10^7), "", "")) +
  scale_y_productlist(labels = c(p10_scale[1:3], "", expression(10^2*"-"*10^7), "", "", "")) +
  labs(x = "OBIS records", y = element_blank(), tag = "A") +
  theme_minimal() +
  theme(legend.position = c(0, 0.5),
        legend.justification = "right") +
  coord_fixed()
)
```
  
Create a plot for the subset of species with a large number (>100) of OBIS records:
```{r}
(obis_gen_hi_obis <- ggplot(data = filter(worms_valid,
                                          obis_cat %in% c("1000", "10000", "1e+05", "1e+06"))) +
  geom_mosaic(aes(x = product(obis_cat), fill = gen_cat), offset = 0.001) +
  scale_fill_viridis_d(name = "GenBank\nnucleotides",
                       labels = p10_scale) +
  scale_x_productlist(labels = c("", "", "", "", p10_scale[5:8])) +
  labs(x = "OBIS records", y = element_blank(), tag = "B") +
  theme_minimal() +
  theme(legend.position = c(0, 0.5), legend.justification = "right",
        axis.text.y = element_blank()) +
  coord_fixed()
  )
```
  
And the same for species with a large number of GenBank nucleotides:
```{r}
(obis_gen_hi_gen <- ggplot(data = filter(worms_valid,
                                         gen_cat %in% c("1000", "10000", "1e+05", "1e+06"))) +
  geom_mosaic(aes(x = product(gen_cat), fill = obis_cat), offset = 0.001) +
  scale_fill_viridis_d(name = "OBIS\nrecords",
                       labels = p10_scale) +
  scale_x_productlist(labels = c("", "", "", "", p10_scale[5:8])) +
  labs(x = "GenBank nucleotides", y = element_blank(), tag = "C") +
  theme_minimal() +
    theme(legend.position = c(0, 0.5), legend.justification = "right",
          axis.text.y = element_blank()) +
  coord_fixed()
)
```
  
The final version of figure 3 was assembled and annotated offline from these three constituent panels, saved here:
```{r}
ggsave("figures/webb_philtrans_fig3a.pdf", obis_gen_all,
       width = 7, height = 6, units = "in")
ggsave("figures/webb_philtrans_fig3b.pdf", obis_gen_hi_obis,
       width = 5.5, height = 4.7, units = "in")
ggsave("figures/webb_philtrans_fig3c.pdf", obis_gen_hi_gen,
       width = 5.5, height = 4.7, units = "in")
```
## IUCN and BOLD data availability
Finally consider IUCN assessments and Barcode of Life data availability. Create a palette for IUCN categories:
```{r}
iucn_palette <- tibble(
  iucn_cat = c("Not Assessed", "Data Deficient", "Threatened", "Not Threatened"),
  iucn_pal = c("#E3DECA", "#077893", "#71AB7F", "#E34F33")
)
```
And plot OBIS records against IUCN categories:
```{r}
iucn_n_obis <- ggplot(filter(worms_valid, in_obis == 1), aes(x = IUCN_cat, y = log10(OBIS),
                                                             colour = IUCN_cat)) +
  geom_quasirandom(size = 0.5, alpha = 3/5) +
  scale_colour_manual(name = "IUCN category",
                    values = pull(iucn_palette, iucn_pal),
                    labels = pull(iucn_palette, iucn_cat),
                    drop = FALSE) +
  labs(x = "IUCN Assessment Status", y = expression(log[10]*"(OBIS records)"), tag = "A") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1),
        legend.position = "none") +
  facet_wrap(~functional_group, ncol = 4, nrow = 2)
```

Do the same for BOLD:
```{r}
bold_n_obis <- ggplot(filter(worms_valid, in_obis == 1), aes(x = in_bold, y = log10(OBIS),
                                                              colour = in_bold)) +
    geom_quasirandom(size = 0.5, alpha = 3/5) +
    scale_colour_manual(name = "In BOLD",
                        values = c("#077893", "#FF9465")) +
    labs(x = "In Barcode of Life Database", y = expression(log[10]*"(OBIS records)"), tag = "B") +
    theme_minimal() +
    theme(legend.position = "none") +
    facet_wrap(~functional_group, ncol = 4, nrow = 2)
```
Assmeble the figure and save to file:
```{r}
(bold_iucn_obis <- iucn_n_obis / bold_n_obis)
ggsave("figures/webb_phitrans_fig4.pdf", bold_iucn_obis, width = 10, height = 10)
```




## Reproducibility
<details><summary>Reproducibility receipt</summary>
```{r}
## datetime
Sys.time()

## repository
git2r::repository()

## session info
sessionInfo()
```

